Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-01-31 PAPER.tex
cleaned up missing section references

@@

This paper is organized into two main sections. In Section \ref{sec:Themes} we introduce the three themes of our focus, using a 
toy model to develop intuition for each one. In Section \ref{sec:CaseStudy} we present a case study into each theme using data 
from the PAPER-64 array, highlighting key changes from the previously published result in \citet{ali_et_al2015}, henceforth known as A15, which have led to a 
revised PAPER-64 power spectrum result (Kolopanis et al., \textit{in prep}). We conclude in Section \ref{sec:Con}.

@@

Signal loss can arise in a variety of ways in the analysis pipeline, such as fitting a polynomial during 
spectral calibration, applying a delay-domain filter, or by deriving weights from data and applying them to itself. Here we focus on signal loss associated with 
applying a weighting matrix to data, a loss that can be significant depending on the choice of weighting and one that was 
previously underestimated in the A15 analysis.

@@

The danger of an empirically estimated covariance matrix comes mostly from not being able to describe the `weak' eigenmodes of $
\textbf{C}$ accurately, for which the EoR signal is brighter than foregrounds. In such a case, the coupling between these modes to the data realization leads to the overfitting and subtraction of the EoR signal. More specifically, the coupling between the estimated covariance and the data is anti-correlated in nature (which is explained in more detail in Section \ref{sec:Illustration}), which leads to loss. Mis-estimating $\textbf{C}$ for EoR-dominated eigenmodes is therefore more harmful than for FG-dominated modes, and since the `weak' modes of an eigenspectrum are typically EoR-dominated, using this part of the spectrum for weighting is most dangerous.

@@

For a particular 
baseline length, there is a maximum delay imposed on sources attached to the sky, which corresponds to the light-crossing time between two 
antennas in a baseline. For longer baselines, this value increases, producing what is known as ``the 
wedge"
\citep{datta_et_al2010, parsons_et_al2012b, vedantham_et_al2012, pober_et_al2013, thyagarajan_et_al2013, liu_et_al2014a, liu_et_al2014b, patil_et_al2017}. 
The wedge describes a region in $k$-space contaminated by smooth spectrum foregrounds, bounded by baseline length (which is proportional to $k_{\perp}$) and delay (which is 
proportional to $k_{\parallel}$). Properties of the wedge can be used to isolate and 
%remove 
avoid foregrounds, as done by A15, 
\citet{parsons_et_al2014}, \citet{dillon_et_al2014}, \citet{dillon_et_al2015}, \citet{jacobs_et_al2015}, \citet{beardsley_et_al2016}, and \citet{trott_et_al2016}.

@@

There are analysis techniques to mitigate the effects of foreground leakage and prevent information from low $k's$ from 
spreading to high $k$ values. For example, narrow window functions in delay-space can be used to minimize the leakage from a particular 
$k$ value into other ones (\citealt{liu_et_al2014b}). In other words, one can construct an estimator using OQE that forces a 
window function to have a minimum response to low $k$ values. The window function used in A15 is constructed in such a way, 
specifically to prevent foregrounds that live at low $k's$ from contaminating higher $k$-modes (see Section \ref{sec:Bias}).

Determining the source of positive non-EoR detections at higher $k$'s is more difficult. In previous power spectrum results, these detections have been explained as instrumental systematics, particularly time-variable cross talk, RFI, cable reflections, and calibration errors (A15; \citealt{parsons_et_al2014}; \citealt{dillon_et_al2014}; \citealt{beardsley_et_al2016}; \citealt{patil_et_al2017}). In the next section (Section \ref{sec:JackknifeOverview}), we will present some tests that can help distinguish these excesses from that of EoR.

@@

We now apply these lessons to data from the PAPER experiment to make a new analysis of the PAPER-64 dataset originally presented in 
A15 and obtain a revised power spectrum estimate.

As a brief review, PAPER is a dedicated $21$ cm experiment located in the Karoo Desert in South Africa. The PAPER-64 
configuration consists of 64 dual-polarization drift-scan elements that are arranged in a grid layout. For our case study, we 
focus solely on Stokes I estimated data \citep{moore_et_al2013} from PAPER's $30$ m East/West baselines (Figure 
\ref{fig:ant_layout}). All data is compressed, calibrated (using self-calibration and redundant calibration), delay-filtered (to remove foregrounds inside the wedge), LST-binned, and fringe-rate filtered. For detailed information about the backend system of PAPER-64, its observations, and data reduction pipeline, we 
refer the reader to \citet{parsons_et_al2010} and A15. We note that all data processing steps are identical to those in A15 until after the LST-binning step in Figure 3 of A15.

@@

The previously best published $21$ cm upper limit result from A15 uses $124$ nights of data to place a $2\sigma$ upper limit 
on $\Delta^{2}(k)$, defined as

@@

For our analysis, we use $8.1$ hours of LST (RA $0.5$-$8.6$ hours) and $51$ total baselines (A15 uses a slightly different RA 
range of $0$-$8.6$ hours). All power spectrum results are produced for a center frequency of $151$ MHz using a width of $10$ 
MHz ($20$ channels), identical to the analysis in A15. We note that, besides using only one baseline type instead of the three as in 
A15, the PAPER-64 dataset that we use in this case study differs from that in A15 only in the analysis. The most significant changes occur in the power spectrum analysis, which is explained in the rest of this paper, but we also note that the applied fringe-rate filter is also slightly different. In A15, the 
applied filter was degraded by widening it in fringe-rate space. This was chosen in order to increase the number of independent 
modes and reduce signal loss, though as we will explain in the next section, this signal loss was still under-estimated. With the development of a new, 
robust method for assessing signal loss, we choose to use the optimal filter in order to maximize sensitivity. This filter is 
computed for a fiducial $30$ m baseline at $150$ MHz, the center frequency in our band. The filter in both the fringe-rate 
domain and time domain is shown in Figure \ref{fig:frp}.

@@

In Section \ref{sec:SiglossOverview}, we showed how signal loss arises when weighting data using information from the data itself. Here we describe a methodology for estimating the amount of signal loss caused by a particular power spectrum estimator when applied to a particular dataset. The exact amount of signal loss will depend on the specific realizations of the signals present in the data and is not something we can directly compute. In this work, as in A15, we inject simulated cosmological signals into our data and test the recovery of those signals (an approach also taken by \citet{masui_et_al2013}). As we will see, correlations between the injected signals and the data are significant complicating factors which were previously not taken into account.

The results of such an injection scheme can be interpreted in multiple ways. In this paper we describe two --- this section describes one and in Appendix \ref{sec:Pr_appendix} we present an alternate interpretation. These two methods yield very similar results, but fundamentally are answering subtly distinct questions. We also highlight major differences from the signal loss computation used in A15, which previously underestimated losses.

\subsubsection{Signal Loss Methodology} 
\label{sec:siglossmethod}
In short, our method consists of injecting an EoR-like signal into the data and then measuring the amount recovered, with the difference being accounted as signal loss.  To capture the full statistical likelihood of signal loss, one requires a quick way to generate many realizations of simulated $21$ 
cm signal visibilities. Here we use the same method as in A15, where mock Gaussian noise visibilities (mock EoR signals) 
are filtered in fringe-rate space to retain only ``sky-like" time-modes. This signal is then added to the data.\footnote{One 
specific change from A15 is that we add this simulated signal into the analysis pipeline before the final fringe-rate filter is 
applied to the data. Previously, the addition was done after that final fringe-rate filter step.  This change results in an increased 
estimate of signal loss, %(by a factor of $\sim$$10$), 
likely due to the use of the fringe-rate filter as a simulator. However, this pipeline difference, while significant, is not the dominant reason why signal loss was under-estimated in A15 (the dominant reason is explained in the main text in Section \ref{sec:siglossmethod}).}

@@

\begin{eqnarray}
\label{eq:crossterm}
P_{\rm out,\alpha} &=& \text{M}^{\alpha}_{r}(\textbf{x}+\textbf{e})^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}(\textbf{x}+\textbf{e}) - 
\text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x} \nonumber \\
&=& \text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \text{M}^{\alpha}_{b}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}
^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& \text{M}^{\alpha}_{c}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} + \text{M}^{\alpha}_{d}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} \nonumber \\
&-& \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}
^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}
Assuming \textbf{R}$_{r}$ is symmetric, the two cross-terms can be summed together as:

@@

{\bf In between:}
When the injected signal is of a similar amplitude to the data by itself, the situation becomes less straightforward. We see that 
the weighted injected power spectrum component mirrors the input power indicating little loss (i.e. the green curve follows the dotted black line), eventually 
departing from unity when the injected amplitude is well above the level of the data power spectrum. However, 
the cross-terms have nearly the same amplitude, but with a negative sign. The sum of the two is substantially lower and therefore drives down the $P_{\rm out}$ estimator (black). We note that in A15, signal loss was computed by only looking at the second term in Equation \eqref{eq:crossterm} (green), which incorrectly implies no loss at the data level. Ignoring the effect of the negative power from the cross-terms is the main reason for under-estimating power spectrum limits in A15.

%A brief description of Figure \ref{fig:sigloss_terms} is as follows. The gray curve represents the data level, which is constant as the level of $P_{\rm in}$, the input EoR signal, is changed. The blue curve is the power spectrum of the data only, weighted by an empirical covariance estimated from the data plus the EoR signal ($\widehat{\textbf{C}}_{r}^{-1}$). It approaches the gray at low injection level, as expected, since in that case $\widehat{\textbf{C}}_{r}^{-1} \rightarrow \widehat{\textbf{C}}_{x}^{-1}$. The green curve is the power spectrum of the EoR signal, weighted by $\widehat{\textbf{C}}_{r}^{-1}$. This is the exact term that was compared to $P_{\rm in}$ in the A15 signal loss analysis, and in doing so this would imply there is negligible loss (where the green crosses the gray). It is primarily due to the cross-terms (red), that the black curve exhibits $\sim$$3$ orders of magnitude of loss at the data level, for the inverse covariance weighted case. We see that the cross-terms can produce a large, negative signal which serves to reduce $P_{\rm out}$ and drive the bulk of our loss. In the unweighted case, however (right plot), the cross-terms primarily drive a tail of `excess' $P_{\rm out}$ at low injects, though at larger injects we see unity transfer as expected (the solid black curve lies on the dotted black line, which represents unity transfer).

%\dcj{This section isn't held up by figure 17. Is that because the "illustriative" figure is off? Or is it that cross terms don't matter when injected signal is very large.}
%We now turn to building intuition behind the negative power that arises from the cross-terms in Equation \eqref{eq:crossterm}. For illustrative purposes in this subsection only, we consider a scenario where $\textbf{e}$ dominates compared to $\textbf{x}$. It is in this limit where we expect signal loss to be greatest, and $P_{\rm out}$ becomes:
%
%\begin{eqnarray}
%\label{eq:pout_expand}
%P_{\rm out, \alpha} &\approx&  M^{\alpha}_{b}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} +
% 2M^{\alpha}_{c}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e}.
%\end{eqnarray}
%The first term of $P_{\rm out}$ is simply the weighted power spectrum of $\textbf{e}$ alone. Again, it is this quantity that was used in A15 to calculate signal loss. However, without taking into account the cross-terms (the second term), we can substantially under-estimate loss. In short, this is because correlations between the data and injected EoR signal emerge as statistical noise that is added to the power spectrum estimate, decreasing the value of $P_{\rm out}$. Including these cross-terms in our new analysis represents the bulk change of our revised analysis compared to A15.

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_terms.png}
	\caption{Illustration of the components of the signal loss transfer function, $P_{\rm out}$, as defined by Equation \eqref{eq:crossterm} for PAPER-64 data for the inverse covariance weighted case (left) and unweighted case (right). The details of the simulation used to generate the figure is explained in Section \ref{sec:Practice}; here we fit smooth polynomials to our data points to create the curves. We show the contributions of each signal loss term (blue, red, green, gray), which when combined produce the black $P_{\rm out}$ curve. Most notably, the green curves denotes the signal loss analysis used in A15, and the red curves showcase the importance of the cross-terms in Equation \eqref{eq:crossterm} in computing signal loss. As explained in the main text, the cross-terms can produce large, negative signal which represents the difference between the green curve (which exhibits negligible loss at the data level) and the black curve (which exhibits $\sim$$3$ orders of magnitude of loss at the data level for the inverse covariance weighted case). Finally, the dotted diagonal black line represents perfect unity transfer, or no signal loss, which we recover for the unweighted case.}
	\label{fig:sigloss_terms}
\end{figure*}

@@

We now shift our attention towards computing signal loss for the fringe-rate filtered PAPER-64 dataset. While our methodology 
outlined below is independent of weighting scheme, here we demonstrate the computation using inverse covariance weighting 
($\textbf{R} = \widehat{\textbf{C}}^{-1}$), the weighting scheme used in A15 which, with an empirically estimated covariance, leads to substantial loss. With this weighting, our 
expressions for $P_{\rm in}$ and $P_{\rm out}$ become

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_transfercurve_posneg.png}
	\caption{Signal loss transfer functions showing the relationship of $P_{\rm in}$ and $P_{\rm out}$, as defined by Equations \eqref{eq:Pin} and \eqref{eq:sigloss}. Kernel density estimations of the power spectrum transfer functions are shown as colored 
heat-maps for the cases of inverse covariance weighted PAPER-64 data (left) and unweighted data (right). The dotted black 
diagonal lines mark a perfect unity mapping, and the solid gray horizontal line denotes the power spectrum value of the data $\widehat{\textbf{p}}_{x}$. From these plots, it is clear that inverse covariance weighting results in $\sim3$ orders of magnitude of signal loss 
for power spectrum values above $\sim$$2 \times 10^{4}$ $mK^{2} \, (h^{-1} \, Mpc)^{3}$, whereas the unweighted case does 
not exhibit loss. The peculiar `signal gain' tail seen at low injection is driven by the cross-terms in Equation \eqref{eq:pout_expand}, also shown by the red curve in the right panel in Figure \ref{fig:sigloss_terms}.}
	\label{fig:sigloss_transfercurve}
\end{figure*}

@@

Our revised, best to-date PAPER-64 power spectrum (using only one baseline separation type and $\widehat{\textbf{C}}_{eff}$) is shown in Figure 
\ref{fig:ps1_data}. Again, black and gray points correspond to positive and negative power spectrum values respectively, with 
$2\sigma$ errors bars. Also plotted are the unweighted power spectrum upper limit (dashed blue) and theoretical prediction of 
noise (solid green). From this result, we quote a best $2\sigma$ upper limit of $(151.0$ mK$)^{2}$ at $k=0.34$ $h$Mpc$^{-1}$, 
a higher limit than A15 by a factor of $\sim7$ in mK (though we only use one baseline type in our analysis as opposed to three). 
\cc{replace with kolopanis limit instead later}

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_data_add.png}
	\includegraphics[width=0.4\textwidth]{plots/ps2_data_add.png}
	\caption{Power spectrum of PAPER-64 using $\widehat{\textbf{C}}_{eff}$. Black and gray points correspond to positive and 
negative power spectrum values, respectively, with $2\sigma$ error bars also plotted. The dashed blue line is the unweighted 
power spectrum ($2\sigma$ upper limit). The solid green line is the theoretical $2\sigma$ noise level prediction based on 
observational parameters. This power spectrum result differs from A15 in that it only uses data from one type of baseline ($30$ m 
East/West baselines) instead of three. Major differences from previously published results stem from revisions regarding signal 
loss, bootstrapping, and the theoretical error computation. From this, we quote a best $2\sigma$ upper limit of $(151.0$ mK$)^{2}$ at 
$k=0.34$ $h$Mpc$^{-1}$, a higher limit than A15 by a factor of $\sim7$ in mK \cc{replace with kolopanis limit later}.}
	\label{fig:ps1_data}
\end{figure*}

@@

In particular, we highlight major changes in both our sensitivity calculation and bootstrapping method that differ from the A15 
analysis of PAPER-64. While we do not discuss the changes within the context of PAPER-32, it is worth noting that the power 
spectrum results in \citet{parsons_et_al2014} are affected by the same issues.

@@

Our revised sensitivity estimate for PAPER-64 is shown in comparison with that of A15 in Figure \ref{fig:sense_check}. 
Together, the revised parameters yield a decrease in sensitivity (higher noise floor) by a factor of $\sim5$ in mK$^{2}$.

>>>>

\defcitealias{ali_et_al2015}{A15}

@@

This paper is organized into two main sections. In Section \ref{sec:Themes} we introduce the three themes of our focus, using a 
toy model to develop intuition for each one. In Section \ref{sec:CaseStudy} we present a case study into each theme using data 
from the PAPER-64 array, highlighting key changes from the previously published result in \citet{ali_et_al2015}, henceforth known as \citetalias{ali_et_al2015}, which have led to a 
revised PAPER-64 power spectrum result (Kolopanis et al., \textit{in prep}). We conclude in Section \ref{sec:Con}.

@@

Signal loss can arise in a variety of ways in the analysis pipeline, such as fitting a polynomial during 
spectral calibration, applying a delay-domain filter, or by deriving weights from data and applying them to itself. Here we focus on signal loss associated with 
applying a weighting matrix to data, a loss that can be significant depending on the choice of weighting and one that was 
previously underestimated in the \citetalias{ali_et_al2015} analysis.

@@

The danger of an empirically estimated covariance matrix comes mostly from not being able to describe the `weak' eigenmodes of $
\textbf{C}$ accurately, for which the EoR signal is brighter than foregrounds. In such a case, the coupling between these modes to the data realization leads to the overfitting and subtraction of the EoR signal. More specifically, the coupling between the estimated covariance and the data is anti-correlated in nature (which is explained in more detail in Section \ref{sec:siglossmethod}), which leads to loss. Mis-estimating $\textbf{C}$ for EoR-dominated eigenmodes is therefore more harmful than for FG-dominated modes, and since the `weak' modes of an eigenspectrum are typically EoR-dominated, using this part of the spectrum for weighting is most dangerous.

@@

For a particular 
baseline length, there is a maximum delay imposed on sources attached to the sky, which corresponds to the light-crossing time between two 
antennas in a baseline. For longer baselines, this value increases, producing what is known as ``the 
wedge"
\citep{datta_et_al2010, parsons_et_al2012b, vedantham_et_al2012, pober_et_al2013, thyagarajan_et_al2013, liu_et_al2014a, liu_et_al2014b, patil_et_al2017}. 
The wedge describes a region in $k$-space contaminated by smooth spectrum foregrounds, bounded by baseline length (which is proportional to $k_{\perp}$) and delay (which is 
proportional to $k_{\parallel}$). Properties of the wedge can be used to isolate and 
%remove 
avoid foregrounds, as done by \citetalias{ali_et_al2015}, 
\citet{parsons_et_al2014}, \citet{dillon_et_al2014}, \citet{dillon_et_al2015}, \citet{jacobs_et_al2015}, \citet{beardsley_et_al2016}, and \citet{trott_et_al2016}.

@@

There are analysis techniques to mitigate the effects of foreground leakage and prevent information from low $k's$ from 
spreading to high $k$ values. For example, narrow window functions in delay-space can be used to minimize the leakage from a particular 
$k$ value into other ones (\citealt{liu_et_al2014b}). In other words, one can construct an estimator using OQE that forces a 
window function to have a minimum response to low $k$ values. The window function used in \citetalias{ali_et_al2015} is constructed in such a way, 
specifically to prevent foregrounds that live at low $k's$ from contaminating higher $k$-modes (see Section \ref{sec:Bias}).

Determining the source of positive non-EoR detections at higher $k$'s is more difficult. In previous power spectrum results, these detections have been explained as instrumental systematics, particularly time-variable cross talk, RFI, cable reflections, and calibration errors (\citetalias{ali_et_al2015}; \citealt{parsons_et_al2014}; \citealt{dillon_et_al2014}; \citealt{beardsley_et_al2016}; \citealt{patil_et_al2017}). In the next section (Section \ref{sec:JackknifeOverview}), we will present some tests that can help distinguish these excesses from that of EoR.

@@

We now apply these lessons to data from the PAPER experiment to make a new analysis of the PAPER-64 dataset originally presented in 
\citetalias{ali_et_al2015} and obtain a revised power spectrum estimate.

As a brief review, PAPER is a dedicated $21$ cm experiment located in the Karoo Desert in South Africa. The PAPER-64 
configuration consists of 64 dual-polarization drift-scan elements that are arranged in a grid layout. For our case study, we 
focus solely on Stokes I estimated data \citep{moore_et_al2013} from PAPER's $30$ m East/West baselines (Figure 
\ref{fig:ant_layout}). All data is compressed, calibrated (using self-calibration and redundant calibration), delay-filtered (to remove foregrounds inside the wedge), LST-binned, and fringe-rate filtered. For detailed information about the backend system of PAPER-64, its observations, and data reduction pipeline, we 
refer the reader to \citet{parsons_et_al2010} and \citetalias{ali_et_al2015}. We note that all data processing steps are identical to those in \citetalias{ali_et_al2015} until after the LST-binning step in Figure 3 of \citetalias{ali_et_al2015}.

@@

The previously best published $21$ cm upper limit result from \citetalias{ali_et_al2015} uses $124$ nights of data to place a $2\sigma$ upper limit 
on $\Delta^{2}(k)$, defined as

@@

For our analysis, we use $8.1$ hours of LST (RA $0.5$-$8.6$ hours) and $51$ total baselines (\citetalias{ali_et_al2015} uses a slightly different RA 
range of $0$-$8.6$ hours). All power spectrum results are produced for a center frequency of $151$ MHz using a width of $10$ 
MHz ($20$ channels), identical to the analysis in \citetalias{ali_et_al2015}. We note that, besides using only one baseline type instead of the three as in 
\citetalias{ali_et_al2015}, the PAPER-64 dataset that we use in this case study differs from that in \citetalias{ali_et_al2015} only in the analysis. The most significant changes occur in the power spectrum analysis, which is explained in the rest of this paper, but we also note that the applied fringe-rate filter is also slightly different. In \citetalias{ali_et_al2015}, the 
applied filter was degraded by widening it in fringe-rate space. This was chosen in order to increase the number of independent 
modes and reduce signal loss, though as we will explain in the next section, this signal loss was still under-estimated. With the development of a new, 
robust method for assessing signal loss, we choose to use the optimal filter in order to maximize sensitivity. This filter is 
computed for a fiducial $30$ m baseline at $150$ MHz, the center frequency in our band. The filter in both the fringe-rate 
domain and time domain is shown in Figure \ref{fig:frp}.

@@

In Section \ref{sec:SiglossOverview}, we showed how signal loss arises when weighting data using information from the data itself. Here we describe a methodology for estimating the amount of signal loss caused by a particular power spectrum estimator when applied to a particular dataset. The exact amount of signal loss will depend on the specific realizations of the signals present in the data and is not something we can directly compute. In this work, as in \citetalias{ali_et_al2015}, we inject simulated cosmological signals into our data and test the recovery of those signals (an approach also taken by \citet{masui_et_al2013}). As we will see, correlations between the injected signals and the data are significant complicating factors which were previously not taken into account.

The results of such an injection scheme can be interpreted in multiple ways. In this paper we describe two --- this section describes one and in Appendix \ref{sec:Pr_appendix} we present an alternate interpretation. These two methods yield very similar results, but fundamentally are answering subtly distinct questions. We also highlight major differences from the signal loss computation used in \citetalias{ali_et_al2015}, which previously underestimated losses.

\subsubsection{Signal Loss Methodology} 
\label{sec:siglossmethod}
In short, our method consists of injecting an EoR-like signal into the data and then measuring the amount recovered, with the difference being accounted as signal loss.  To capture the full statistical likelihood of signal loss, one requires a quick way to generate many realizations of simulated $21$ 
cm signal visibilities. Here we use the same method as in \citetalias{ali_et_al2015}, where mock Gaussian noise visibilities (mock EoR signals) 
are filtered in fringe-rate space to retain only ``sky-like" time-modes. This signal is then added to the data.\footnote{One 
specific change from \citetalias{ali_et_al2015} is that we add this simulated signal into the analysis pipeline before the final fringe-rate filter is 
applied to the data. Previously, the addition was done after that final fringe-rate filter step.  This change results in an increased 
estimate of signal loss, %(by a factor of $\sim$$10$), 
likely due to the use of the fringe-rate filter as a simulator. However, this pipeline difference, while significant, is not the dominant reason why signal loss was under-estimated in \citetalias{ali_et_al2015} (the dominant reason is explained in the main text in Section \ref{sec:siglossmethod}).}

@@

\begin{eqnarray}
\label{eq:crossterm_full}
P_{\rm out,\alpha} &=& \text{M}^{\alpha}_{r}(\textbf{x}+\textbf{e})^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}(\textbf{x}+\textbf{e}) - 
\text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x} \nonumber \\
&=& \text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \text{M}^{\alpha}_{b}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}
^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& \text{M}^{\alpha}_{c}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} + \text{M}^{\alpha}_{d}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} \nonumber \\
&-& \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}
^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}
Assuming \textbf{R}$_{r}$ is symmetric, the two cross-terms can be summed together as:

@@

{\bf In between:}
When the injected signal is of a similar amplitude to the data by itself, the situation becomes less straightforward. We see that 
the weighted injected power spectrum component mirrors the input power indicating little loss (i.e. the green curve follows the dotted black line), eventually 
departing from unity when the injected amplitude is well above the level of the data power spectrum. However, 
the cross-terms have nearly the same amplitude, but with a negative sign. The sum of the two is substantially lower and therefore drives down the $P_{\rm out}$ estimator (black). We note that in \citetalias{ali_et_al2015}, signal loss was computed by only looking at the second term in Equation \eqref{eq:crossterm} (green), which incorrectly implies no loss at the data level. Ignoring the effect of the negative power from the cross-terms is the main reason for under-estimating power spectrum limits in \citetalias{ali_et_al2015}.

%A brief description of Figure \ref{fig:sigloss_terms} is as follows. The gray curve represents the data level, which is constant as the level of $P_{\rm in}$, the input EoR signal, is changed. The blue curve is the power spectrum of the data only, weighted by an empirical covariance estimated from the data plus the EoR signal ($\widehat{\textbf{C}}_{r}^{-1}$). It approaches the gray at low injection level, as expected, since in that case $\widehat{\textbf{C}}_{r}^{-1} \rightarrow \widehat{\textbf{C}}_{x}^{-1}$. The green curve is the power spectrum of the EoR signal, weighted by $\widehat{\textbf{C}}_{r}^{-1}$. This is the exact term that was compared to $P_{\rm in}$ in the \citetalias{ali_et_al2015} signal loss analysis, and in doing so this would imply there is negligible loss (where the green crosses the gray). It is primarily due to the cross-terms (red), that the black curve exhibits $\sim$$3$ orders of magnitude of loss at the data level, for the inverse covariance weighted case. We see that the cross-terms can produce a large, negative signal which serves to reduce $P_{\rm out}$ and drive the bulk of our loss. In the unweighted case, however (right plot), the cross-terms primarily drive a tail of `excess' $P_{\rm out}$ at low injects, though at larger injects we see unity transfer as expected (the solid black curve lies on the dotted black line, which represents unity transfer).

%\dcj{This section isn't held up by figure 17. Is that because the "illustriative" figure is off? Or is it that cross terms don't matter when injected signal is very large.}
%We now turn to building intuition behind the negative power that arises from the cross-terms in Equation \eqref{eq:crossterm}. For illustrative purposes in this subsection only, we consider a scenario where $\textbf{e}$ dominates compared to $\textbf{x}$. It is in this limit where we expect signal loss to be greatest, and $P_{\rm out}$ becomes:
%
%\begin{eqnarray}
%\label{eq:pout_expand}
%P_{\rm out, \alpha} &\approx&  M^{\alpha}_{b}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} +
% 2M^{\alpha}_{c}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e}.
%\end{eqnarray}
%The first term of $P_{\rm out}$ is simply the weighted power spectrum of $\textbf{e}$ alone. Again, it is this quantity that was used in \citetalias{ali_et_al2015} to calculate signal loss. However, without taking into account the cross-terms (the second term), we can substantially under-estimate loss. In short, this is because correlations between the data and injected EoR signal emerge as statistical noise that is added to the power spectrum estimate, decreasing the value of $P_{\rm out}$. Including these cross-terms in our new analysis represents the bulk change of our revised analysis compared to \citetalias{ali_et_al2015}.

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_terms.png}
	\caption{Illustration of the components of the signal loss transfer function, $P_{\rm out}$, as defined by Equation \eqref{eq:crossterm} for PAPER-64 data for the inverse covariance weighted case (left) and unweighted case (right). The details of the simulation used to generate the figure is explained in Section \ref{sec:Practice}; here we fit smooth polynomials to our data points to create the curves. We show the contributions of each signal loss term (blue, red, green, gray), which when combined produce the black $P_{\rm out}$ curve. Most notably, the green curves denotes the signal loss analysis used in \citetalias{ali_et_al2015}, and the red curves showcase the importance of the cross-terms in Equation \eqref{eq:crossterm} in computing signal loss. As explained in the main text, the cross-terms can produce large, negative signal which represents the difference between the green curve (which exhibits negligible loss at the data level) and the black curve (which exhibits $\sim$$3$ orders of magnitude of loss at the data level for the inverse covariance weighted case). Finally, the dotted diagonal black line represents perfect unity transfer, or no signal loss, which we recover for the unweighted case.}
	\label{fig:sigloss_terms}
\end{figure*}

@@

We now shift our attention towards computing signal loss for the fringe-rate filtered PAPER-64 dataset. While our methodology 
outlined below is independent of weighting scheme, here we demonstrate the computation using inverse covariance weighting 
($\textbf{R} = \widehat{\textbf{C}}^{-1}$), the weighting scheme used in \citetalias{ali_et_al2015} which, with an empirically estimated covariance, leads to substantial loss. With this weighting, our 
expressions for $P_{\rm in}$ and $P_{\rm out}$ become

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_transfercurve_posneg.png}
	\caption{Signal loss transfer functions showing the relationship of $P_{\rm in}$ and $P_{\rm out}$, as defined by Equations \eqref{eq:Pin} and \eqref{eq:sigloss}. Kernel density estimations of the power spectrum transfer functions are shown as colored 
heat-maps for the cases of inverse covariance weighted PAPER-64 data (left) and unweighted data (right). The dotted black 
diagonal lines mark a perfect unity mapping, and the solid gray horizontal line denotes the power spectrum value of the data $\widehat{\textbf{p}}_{x}$. From these plots, it is clear that inverse covariance weighting results in $\sim3$ orders of magnitude of signal loss 
for power spectrum values above $\sim$$2 \times 10^{4}$ $mK^{2} \, (h^{-1} \, Mpc)^{3}$, whereas the unweighted case does 
not exhibit loss. The peculiar `signal gain' tail seen at low injection is driven by the cross-terms in Equation \eqref{eq:crossterm_full}, also shown by the red curve in the right panel in Figure \ref{fig:sigloss_terms}.}
	\label{fig:sigloss_transfercurve}
\end{figure*}

@@

Our revised, best to-date PAPER-64 power spectrum (using only one baseline separation type and $\widehat{\textbf{C}}_{eff}$) is shown in Figure 
\ref{fig:ps1_data}. Again, black and gray points correspond to positive and negative power spectrum values respectively, with 
$2\sigma$ errors bars. Also plotted are the unweighted power spectrum upper limit (dashed blue) and theoretical prediction of 
noise (solid green). From this result, we quote a best $2\sigma$ upper limit of $(151.0$ mK$)^{2}$ at $k=0.34$ $h$Mpc$^{-1}$, 
a higher limit than \citetalias{ali_et_al2015} by a factor of $\sim7$ in mK (though we only use one baseline type in our analysis as opposed to three). 
\cc{replace with kolopanis limit instead later}

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_data_add.png}
	\includegraphics[width=0.4\textwidth]{plots/ps2_data_add.png}
	\caption{Power spectrum of PAPER-64 using $\widehat{\textbf{C}}_{eff}$. Black and gray points correspond to positive and 
negative power spectrum values, respectively, with $2\sigma$ error bars also plotted. The dashed blue line is the unweighted 
power spectrum ($2\sigma$ upper limit). The solid green line is the theoretical $2\sigma$ noise level prediction based on 
observational parameters. This power spectrum result differs from \citetalias{ali_et_al2015} in that it only uses data from one type of baseline ($30$ m 
East/West baselines) instead of three. Major differences from previously published results stem from revisions regarding signal 
loss, bootstrapping, and the theoretical error computation. From this, we quote a best $2\sigma$ upper limit of $(151.0$ mK$)^{2}$ at 
$k=0.34$ $h$Mpc$^{-1}$, a higher limit than \citetalias{ali_et_al2015} by a factor of $\sim7$ in mK \cc{replace with kolopanis limit later}.}
	\label{fig:ps1_data}
\end{figure*}

@@

In particular, we highlight major changes in both our sensitivity calculation and bootstrapping method that differ from the \citetalias{ali_et_al2015} 
analysis of PAPER-64. While we do not discuss the changes within the context of PAPER-32, it is worth noting that the power 
spectrum results in \citet{parsons_et_al2014} are affected by the same issues.

@@

Our revised sensitivity estimate for PAPER-64 is shown in comparison with that of \citetalias{ali_et_al2015} in Figure \ref{fig:sense_check}. 
Together, the revised parameters yield a decrease in sensitivity (higher noise floor) by a factor of $\sim5$ in mK$^{2}$.
